---
title: "Analysis of Bauer et al. (2023) Beta diversity on dike grasslands: <br> Spatial variation 2019"
author: "<b>Markus Bauer</b> <br>"
date: "<b>`r format(Sys.time(), '%Y-%m-%d')`</b>"
output:
  github_document:
    toc: true
    toc_depth: 3
    dev: png
    fig_width: 7
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

<br/>
<br/>
<b>Markus Bauer</b>

Technichal University of Munich, TUM School of Life Sciences, Chair of
Restoration Ecology, Emil-Ramann-Straße 6, 85354 Freising, Germany

[markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

ORCiD ID: [0000-0001-5372-4174](https://orcid.org/0000-0001-5372-4174)
<br>
[Google Scholar](https://scholar.google.de/citations?user=oHhmOkkAAAAJ&hl=de&oi=ao)
<br>
GitHub: [markus1bauer](https://github.com/markus1bauer)

To compare different models, you only have to change the models in
section 'Load models'

# Preparation

Borcard, Gillet & Legendre (2018) Numerical Ecology with R. 2nd edition. Springer, Cham. [DOI: 10.1007/978-3-319-71404-2](https://doi.org/10.1007/978-3-319-71404-2) Chapter 6.3

#### Packages

```{r libraries, message = FALSE}
library(here)
library(tidyverse)
library(vegan)
library(adespatial)
```

```{r echo = FALSE}
rm(list = ls())
```

#### Load data

```{r load-data}
sites <- read_csv(here("data", "processed", "data_processed_sites_spatial.csv"),
  col_names = TRUE,
  na = c("na", "NA"), col_types =
    cols(
      .default = "?",
      id = "f",
      location_abb = "f",
      block = "f",
      plot = "f",
      exposition = "f",
      orientation = "f",
      location_construction_year = "f"
    )) %>%
  filter(survey_year == 2019) %>%
  select(
    id, plot, block, longitude, latitude,
    botanist, location_construction_year, construction_year,
    exposition, orientation, pc1_soil, pc2_soil, pc3_soil,
    location_abb, river_km, river_distance, biotope_distance, biotope_area,
    mem1_2019,
    survey_year, plot_age, pc1_construction_year, pc2_construction_year,
    pc3_construction_year,
    accumulated_cover
    ) %>%
  mutate(
    survey_year_factor = as_factor(survey_year),
    exposition_numeric = as.double(exposition),
    orientation_numeric = as.double(orientation),
    location_abb_numeric = as.double(location_abb),
    botanist_numeric = as.double(as_factor(botanist)),
    biotope_area = if_else(is.na(biotope_area), 0, biotope_area)
  )

species <- read_csv(here("data", "processed", "data_processed_species.csv"),
  col_names = TRUE,
  na = c("na", "NA", ""), col_types =
    cols(
      .default = "d",
      name = "f"
    )) %>%
  mutate(across(where(is.numeric), ~ replace(., is.na(.), 0))) %>%
  pivot_longer(-name, names_to = "id", values_to = "value") %>%
  pivot_wider(id, names_from = "name", values_from = "value") %>%
  semi_join(sites, by = "id") %>%
  arrange(id) %>%
  column_to_rownames("id")

sites <- sites %>%
  column_to_rownames("id")
```

# Statistics

## Calculate beta diversity

### Check collinearity

Exclude r > 0.7
<br>
Dormann et al. 2013 Ecography [DOI: 10.1111/j.1600-0587.2012.07348.x](https://doi.org/10.1111/j.1600-0587.2012.07348.x)

```{r collinearity, warning = FALSE, message = FALSE, fig.width = 18, fig.height = 14}
sites %>%
  select(
    where(is.numeric), -ends_with("numeric"),
    -accumulated_cover, -construction_year, -survey_year,
    ) %>%
  GGally::ggpairs(lower = list(continuous = "smooth_loess"))
```

&rarr; Remove longitude, latitude, biotope_area, mem1_2019, pc3_construction_year

```{r data_selection}
sites_soil <- sites %>%
  select(pc1_soil, pc2_soil, pc3_soil, exposition_numeric, orientation_numeric)
sites_space <- sites %>%
  select(location_abb_numeric, river_distance, river_km, biotope_distance)
sites_history <- sites %>%
  select(plot_age, pc1_construction_year, pc2_construction_year)
```

### Calculate: Baselga presence-absence

```{r beta_calculation}
beta <- beta.div.comp(species, coef = "BS", quant = FALSE)
beta$Note
beta$part
beta_total <- beta$D %>% # = Soerensen dissimilarity
  as.matrix() %>%
  as.data.frame()
beta_substitution <- beta$repl %>% # = Replacement / Simpson dissimilarity
  as.matrix()
beta_subsets <- beta$rich %>% # = Nestedness
  as.matrix()
```

## db-RDA: Replacement component

### Check linear trend in data

m1 <- dbrda(beta_substitution ~ longitude + latitude, data = sites)
anova(m1)
beta_substitution_detrended <- resid(lm(beta_substitution ~ longitude + latitude, data = sites))
&rarr; this trend is captured by river_km

### Full model

```{r full_model}
m1 <- dbrda(
  beta_substitution ~
    pc1_soil + pc2_soil + pc3_soil + exposition + orientation +
    location_abb + river_km + river_distance + biotope_distance +
    plot_age + pc1_construction_year + pc2_construction_year,
  data = sites
  )
anova(m1, permutations = how(nperm = 9999))
(r2adj <- RsquareAdj(m1)$adj.r.squared)
```

### Forward selection: Soil

```{r fs_soil}
m1 <- dbrda(
  beta_substitution ~ pc1_soil + pc2_soil + pc3_soil + exposition + orientation,
  data = sites
  )
r2adj <- RsquareAdj(m1)$adj.r.squared
sel <- forward.sel(
  beta_substitution,
  sites_soil,
  adjR2thresh = r2adj,
  nperm = 9999
  )
sel$p_adj <- p.adjust(sel$pvalue, method = "holm", n = ncol(sites_soil))
sel # https://www.davidzeleny.net/anadat-r/doku.php/en:forward_sel_examples
sites_soil_selected <- sites %>%
  select(exposition_numeric, pc3_soil)
```

### Forward selection: Space

```{r fs_space}
m1 <- dbrda(
  beta_substitution ~ location_abb + river_km + river_distance +
    biotope_distance,
  data = sites
  )
r2adj <- RsquareAdj(m1)$adj.r.squared
#sel <- forward.sel(
#  beta_substitution,
#  sites_space,
#  adjR2thresh = r2adj,
#  nperm = 9999
#  )
```

&rarr; No variables could be selected

### Forward selection: History

```{r fs_history}
m1 <- dbrda(
  beta_substitution ~ plot_age + pc1_construction_year + pc2_construction_year,
  data = sites
  )
r2adj <- RsquareAdj(m1)$adj.r.squared
sel <- forward.sel(
  beta_substitution,
  sites_history,
  adjR2thresh = r2adj,
  nperm = 9999
  )
sel$p_adj <- p.adjust(sel$pvalue, method = "holm", n = ncol(sites_history))
sel # https://www.davidzeleny.net/anadat-r/doku.php/en:forward_sel_examples
sites_history_selected <- sites %>%
  select(pc1_construction_year)
```

### Variation partitioning

```{r varpart, message = FALSE}
m1_substitution_varpart <- varpart(
  beta_substitution, sites_soil_selected, sites_history_selected
)
plot(
  m1_substitution_varpart,
  Xnames = c("Site", "History"),
  cutoff = 0.01, digits = 2, bg = NA
  )
```

```{r varpart_plot, include = FALSE}
m1_substitution_varpart <- varpart(
  beta_substitution, sites_soil_selected, sites_history_selected
)
tiff(
  here("outputs", "figures", "figure_3a_2017_800dpi_12x12cm.tiff"),
  res = 72, width = 12, height = 12, units = "cm", compression = "none"
  )
plot(
  m1_substitution_varpart,
  Xnames = c("Site", "History"),
  cutoff = 0.01, digits = 2, bg = NA
  )
dev.off()
```

### Partial db-RDA

#### Soil

```{r partial_soil}
m1_substitution <- dbrda(
  beta_substitution ~ exposition + pc3_soil +
    Condition(pc1_construction_year),
  data = sites
  )
anova(m1_substitution, permutations = how(nperm = 9999))
RsquareAdj(m1_substitution)
```

#### History = pc1_construction_year

```{r partial_history_pc1-constr-year}
m1_substitution <- dbrda(
  beta_substitution ~ pc1_construction_year +
  Condition(pc3_soil + orientation + exposition + pc1_soil +
    location_abb),
  data = sites
  )
anova(m1_substitution, permutations = how(nperm = 9999))
RsquareAdj(m1_substitution)
```

#### Exposition

```{r partial_exposition}
m1_substitution <- dbrda(
  beta_substitution ~ exposition +
    Condition(pc3_soil + pc1_construction_year),
  data = sites
  )
anova(m1_substitution, permutations = how(nperm = 9999))
RsquareAdj(m1_substitution)
```

#### PC3_soil

```{r partial_pc3-soil}
m1_substitution <- dbrda(
  beta_substitution ~ pc3_soil +
    Condition(exposition + pc1_construction_year),
  data = sites
  )
anova(m1_substitution, permutations = how(nperm = 9999))
RsquareAdj(m1_substitution)
```

## db-RDA: Nestedness component

### Full model

```{r nestedness_full_model}
m1 <- dbrda(
  beta_subsets ~ pc1_soil + pc2_soil + pc3_soil + exposition + orientation +
  location_abb + river_km + river_distance + biotope_distance +
  plot_age + pc1_construction_year + pc2_construction_year,
  data = sites
  )
anova(m1, permutations = how(nperm = 999))
(r2adj <- RsquareAdj(m1)$adj.r.squared)
```

### Forward selection

&rarr; no forward selection because full model is not significant



